\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species

### 2024-04-17 ms.Rnw
Update of gitignor

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear,numbers]{natbib}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyphenat}
\tolerance=1000
\usepackage{url}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{hyphenat}
\usepackage{Sweave}
\bibliographystyle{apalike}
\newcommand{\blind}{1}
\usetikzlibrary{positioning}
\usetikzlibrary {arrows.meta}
\usetikzlibrary{shapes.geometric}
% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\begin{document}
\SweaveOpts{concordance=TRUE}
\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

@@

<<Setup, include=FALSE,echo=FALSE,results=hide>>=
 library(plyr)
 library(reshape2)
library(patchwork)
 library(magrittr)
 library(foreach)
 library(readxl)
 library(doParallel)
 registerDoParallel(cores=36)
 library(pomp)
 library(panelPomp)
 library(ggplot2)
 library(gridExtra)
 library(scales)  # For formatting scales
 library(patchwork)

@@

As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both non-parasitized and prarasitized dynamics to minimize information bias. Figures below illustrates these dynamics, with the first row of Figure 1 depicting non-parasitized dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}. The second row of Figure 1 shows parasitized dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation. Non-parasitized dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species. And based on this result, the correlation between native host and invasive host in parasitized dynamics will be further explored.

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot1", fig=TRUE, width=25, height=10,out.width='\\textwidth'>>=
mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
  cbind(mixed_nopara_data[[i]], Trial = paste("Trial", trails[i]))
}))
mixed_nopara_data_combined$Bucket = paste0(rep("Bucket ",80),rep(1:9, each = 10))

@@

# Adjusting p1 with specified font size modifications
p1 <- ggplot(mixed_nopara_data_combined, aes(x = day, y = sqrt(dent.adult))) +
  geom_line(aes(color = "Dent Adult"), size = 1) +
  geom_line(aes(y = sqrt(lum.adult), color = "Lum Adult"), size = 1) +
  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),  # Y-axis title
        axis.text.x = element_text(size = 10),   # X-axis text
        axis.text.y = element_text(size = 10),   # Y-axis text
        legend.title = element_text(size = 10),  # Legend title
        legend.text = element_text(size = 10),   # Legend text
        strip.text = element_text(size = 10),    # Facet label text
        strip.background = element_blank(),
        strip.placement = "outside") +
  ylab('Square root of Daphnia density') +
  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
  labs(title = "Daphnia Density in Non-parasitized Dynamics", color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue")) +
  theme_minimal() +
  guides(color = guide_legend(title = "Species"))

# Adjusting p2 with specified font size modifications
p2 <- ggplot(mixed_para_data_combined, aes(x = day, y = sqrt(dentadult))) +
  geom_line(aes(color = "Dent Adult"), size = 1) +
  geom_line(aes(y = sqrt(lumadult), color = "Lum Adult"), size = 1) +
  geom_line(aes(y = sqrt(luminf), color = "Lum Inf"), size = 1) +
  geom_line(aes(y = sqrt(dentinf), color = "Dent Inf"), size = 1) +
  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        strip.background = element_blank(),
        strip.placement = "outside") +
  ylab('Square root of Daphnia density') +
  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
  labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red")) +
  theme_minimal() +
  guides(color = guide_legend(title = "Species"))

grid.arrange(p1, p2, ncol = 1)
@
% \includegraphics[width=1\textwidth]{"Density_plot.png"}
% \label{fig:plot1}
\caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}. The Buckets 1 to 9 represents different buckets in different experiment. The black lines represent the \textit{D. dentifera} density, the red lines show the \textit{D. lumholtzi} density, the green lines are the infected \textit{D. dentifera} density, and the blue lines indicate the infected \textit{D. lumholtzi} density.}
\end{figure}

@@

<<Results, include=FALSE,echo=FALSE,results=hide>>=
load('all_shared.RData')
mif.estimate = as.data.frame(t(mif.estimate))

@@

\section{Results}
\label{sec:res}
<<Result2, include=FALSE,echo=FALSE,results=hide>>=
 load('SIRPF_model_estimate.RData')

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[authoryear,numbers]{natbib}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyphenat}
\tolerance=1000
\usepackage{url}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{hyphenat}

\bibliographystyle{apalike}
\newcommand{\blind}{1}
\usetikzlibrary{positioning}
\usetikzlibrary {arrows.meta}
\usetikzlibrary{shapes.geometric}
% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\begin{document}
% \SweaveOpts{concordance=TRUE}
\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

@@

<<Setup, include=FALSE,echo=FALSE,results='hide'>>=
 library(plyr)
 library(reshape2)
library(patchwork)
 library(magrittr)
 library(foreach)
 library(readxl)
 library(doParallel)
 registerDoParallel(cores=36)
 library(pomp)
 library(panelPomp)
 library(ggplot2)
 library(gridExtra)
 library(scales)  # For formatting scales
 library(patchwork)

@@

As juveniles rarely exhibit clear signs of infection, only adult \textit{Daphnia} were used for quantifying effects in both non-parasitized and prarasitized dynamics to minimize information bias. Figures below illustrates these dynamics, with the first row of Figure 1 depicting non-parasitized dynamics (mixed-species) where changes in native and invasive host densities reflect direct resource competition, for example, food alga \textit{Ankistrodesmus falcatus}. The second row of Figure 1 shows parasitized dynamics, where the interplay of susceptible and infected \textit{Daphnia} of different species is further influenced by parasites, complicating the variation of susceptible and infected \textit{Daphnia} of different races, which makes it difficult to directly analyze the sources of influence on their variation. Non-parasitized dynamics will be analyzed at the first place to help understand the effect of resource competition in mixed-species. And based on this result, the correlation between native host and invasive host in parasitized dynamics will be further explored.
% 
\begin{figure}[H]
% \centering
<<echo=FALSE, fig.width=8, fig.height=6,out.width="2in">>=
 plot(x = 1:10)
@
\end{figure}
%

\begin{figure}[H]
\centering
<<DataVis,echo=FALSE, fig.width=9,fig.height=6,out.width="2in">>=
mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
  cbind(mixed_nopara_data[[i]], Trial = paste("Trial", trails[i]))
}))
mixed_nopara_data_combined$Bucket = paste0(rep("Bucket ",80),rep(1:9, each = 10))

@@

# Adjusting p1 with specified font size modifications
p1 <- ggplot(mixed_nopara_data_combined, aes(x = day, y = sqrt(dent.adult))) +
  geom_line(aes(color = "Dent Adult"), size = 1) +
  geom_line(aes(y = sqrt(lum.adult), color = "Lum Adult"), size = 1) +
  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
  theme(axis.title.x = element_blank( ),
        axis.title.y = element_text(size = 10),  # Y-axis title
        axis.text.x = element_text(size = 10),   # X-axis text
        axis.text.y = element_text(size = 10),   # Y-axis text
        legend.title = element_text(size = 10),  # Legend title
        legend.text = element_text(size = 10),   # Legend text
        strip.text = element_text(size = 10),    # Facet label text
        strip.background = element_blank(),
        strip.placement = "outside",
        plot.margin = c(.5,.5,.5,.5)) +
  ylab('Square root of Daphnia density') +
  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
  labs(title = "Daphnia Density in Non-parasitized Dynamics", color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue")) +
  theme_minimal() +
  guides(color = guide_legend(title = "Species"))

# Adjusting p2 with specified font size modifications
p2 <- ggplot(mixed_para_data_combined, aes(x = day, y = sqrt(dentadult))) +
  geom_line(aes(color = "Dent Adult"), size = 1) +
  geom_line(aes(y = sqrt(lumadult), color = "Lum Adult"), size = 1) +
  geom_line(aes(y = sqrt(luminf), color = "Lum Inf"), size = 1) +
  geom_line(aes(y = sqrt(dentinf), color = "Dent Inf"), size = 1) +
  facet_wrap(~Bucket, nrow = 1, labeller = bucket_labeller) +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text = element_text(size = 10),
        strip.background = element_blank(),
        strip.placement = "outside",
        plot.margin = c(.5,.5,.5,.5)) +
  ylab('Square root of Daphnia density') +
  scale_x_continuous(limits = c(0, 52), breaks = c(0, 25, 50)) +
  labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red")) +
  theme_minimal() +
  guides(color = guide_legend(title = "Species"))

grid.arrange(p1, p2, ncol = 1)

@
% \includegraphics[width=1\textwidth]{"Density_plot.png"}
\label{fig:plot1}
\caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}. The Buckets 1 to 9 represents different buckets in different experiment. The black lines represent the \textit{D. dentifera} density, the red lines show the \textit{D. lumholtzi} density, the green lines are the infected \textit{D. dentifera} density, and the blue lines indicate the infected \textit{D. lumholtzi} density.}
\end{figure}

@@

<<Results, include=FALSE,echo=FALSE,results='hide'>>=
load('all_shared.RData')
mif.estimate = as.data.frame(t(mif.estimate))

@@

\section{Results}
\label{sec:res}
<<Result2, include=FALSE,echo=FALSE,results='hide'>>=
 load('SIRPF_model_estimate.RData')
